---
title: "Homework One"
author: "Yi Xiao"
date: 2018-09-11
output: github_document


---

This assignment reinforces ideas in the building blocks topic.


# *Problem 1*
This problem focuses on variable types and coercion 

#### Here's a **code chunk** that create a data frame with:

* A random sample of size 10 from a uniform[0, 5] distribution

* A logical vector indicating whether elements of the sample are greater than 2

* A (length-10) character vector

* A (length-10) factor vector

```{r create_dataframe, warning = FALSE}
set.seed(1)
samp_uni <- runif(10, min = 0, max = 5)
samp_compare <- samp_uni > 2
samp_name <- c("ab","zb","c7y","zd","ab","f","g","h","i","k")
samp_factor <- factor(c(1,2,3,1,2,3,2,1,3,2))
all_data <- data.frame(samp_uni, samp_compare, samp_name, samp_factor)
all_data
```

The **mean** of samp_uni and samp_compare is **`r mean(all_data$samp_uni)`** and **`r mean(all_data$samp_compare)`** respectively, however the mean of other two variables are  **`r mean(all_data$samp_name)`** and  **`r mean(all_data$samp_factor)`** because the argument of the mean function can only be numeric or logical.



#### Here's a **code chunk** that convert logical, character, and factor to numeric:
```{r, results = FALSE, warning = FALSE}
as.numeric(samp_compare)
as.numeric(samp_name)
as.numeric(samp_factor)
```

When convert logical to numeric, TRUE was coverted to 1 and False to 0. While NA was generated by character-to-numerical coercion.  The same factors were converted to the same numericals.

#### Here is a **code chunk** that convert character to factor to numeric:
```{r, warning = FALSE, collapse = TRUE }
# 1. character to factor
as.factor(samp_name)

# 2.character to numeric
as.numeric(samp_name)

# character to factor then to numeric
as.numeric(as.factor(samp_name))
```
When converting character to factor: each element with different name becomes a new level of factor; when converting character to numeric: NA was generated; when converting character to factor then to numeric: numeric was assigned according to alphabetic order.

#### Here's a **code chunk** that convert factor to character to numeric:
```{r, collapse = TRUE}
# 1. factor to character
as.character(samp_factor)
# 2. factor to numerical
as.numeric(samp_factor)
# 3. factor to character then to numerical 
as.numeric(as.character(samp_factor))
```
# *Problem 2*
This problem focuses on plotting and the use of inline R code.

#### Here's a **code chunk** that create a data frame with:

* x: a random sample of size 1000 from a standard Normal distribution

* y: a random sample of size 1000 from a standard Normal distribution

* A logical vector indicating whether the x + y > 0

* A numeric vector created by coercing the above logical vector

* A factor vector created by coercing the above logical vector
```{r}
set.seed(2)
x <- rnorm(1000)
set.seed(3)
y <-rnorm(1000)
sum_greater_zero <- x + y > 0
numer_greater_zero <- as.numeric(sum_greater_zero)
sum_group <- as.factor(numer_greater_zero)
all_data2 <- data.frame(x, y, sum_greater_zero, numer_greater_zero, sum_group)
```

* The size of the dataset: **`r nrow(all_data2)`**

* The mean and median of x: **`r mean(x)`** and **`r median(x)`**

* The proportion of cases for which the logical vector is TRUE:   
**`r sum(all_data2$sum_greater_zero[all_data2$sum_greater_zero == TRUE])/nrow(all_data2)`**


#### Here's a **code chunk** that create a scatterplot:

* *color points using the logical variable*
```{r}
library(ggplot2)
ggplot(all_data2, aes(x = x, y = y)) + geom_point(aes(colour = sum_greater_zero))
ggsave("myplot1.png")
```

* *color points using the numeric variable*
```{r}
library(ggplot2)
ggplot(all_data2, aes(x = x, y = y)) + geom_point(aes(colour = numer_greater_zero))
ggsave("myplot2.png")
```

* *color points using the factor variable*
```{r}
library(ggplot2)
ggplot(all_data2, aes(x = x, y = y)) + geom_point(aes(colour = sum_group))
ggsave("myplot3.png")
```


